### 2014-04-08 reviewer-comments/letter.tex
More letters

\documentclass{letter}
\usepackage{color,soul}
\usepackage{hyperref}
\usepackage{fullpage}
%\usepackage{cleveref}

\address{Clark Center S295 \\ Stanford University \\ Stanford, CA, 94305}
\signature{Robert T. McGibbon, Christian R. Schwantes, Vijay S. Pande}

\newcommand{\separate}{\begin{center}--------\end{center}}

\begin{document}

\begin{letter}{Prof. Sharon Hammes-Schiffer \\ Deputy Editor of JPC B \\ The Journal of Physical Chemistry}
\opening{Editor:}

Reviewer 1
\begin{quote}
In ``Optimal parameter selection in Markov state models for biomolecular conformational dynamics'', the authors elaborate on an earlier likelihood-based approach to the selection of the appropriate number of discrete states for crisp partitionings to attempt to obtain a Markov model that balances accuracy with the potential for overfitting that many-state models are capable of.  Three different critera are examined: a leave-some-out cross-validation likelihood estimate, as well as two criteria that penalize overfitting by using assumptions of asymptotic normality of the likelihood function (AIC and BIC).  Two different "emission probabilities" for the probability of observing a sample $x_t$ from a given discrete state s are considered that would produce a generative model---one that can be used to assign new data to a model: a uniform probability and Gaussian probability.  These are compared for consistency with each other, as well as compared with well-established metrics such as the stability of implied timescales of the model with the number of states.
\end{quote}

@@

In order to compute the Bayes factor for selecting the number of states, we would need to integrate over the space of all crisp state decompositions $S$ with $|S|=k$ as well as the simplex of $k \times k$ reversible stochastic matrices. Although the second is possible analytically using the results of Diaconis and Rolles (2006), we're not aware of any practical methods for the former. To avoid confusion, we've replaced the word ``intractable'' with ``impractical.''

@@

We've expanded our discussion of the simulation methods as well as the parameters used in MSM construction to address these and other issues. The new subsection, entitled \emph{Simulation and MSM Construction} can be found in the \emph{Methods} section. \hl{need to add the initial points}

@@

\begin{quote}
``...the likelihood remained essentially constant...'' What does this mean, quantitatively?
\end{quote}
\hl{It's really just a qualitative change. I just meant that the changes in the likelihood are much less than the differences between models, and so are not important for evaluating the ``best" model.}

\separate
\begin{quote}
according to the three methods" Which three methods exactly?
\end{quote}
We've clarified this to read:
\begin{quote}
according to the BIC, AIC, and test log-likelihood
\end{quote}

\separate
\begin{quote}
Fig. 2: This figure is way too small.  What is ``score''?  Is this a log likelihood?  Is lower or higher score representative of lower or higher likelihood?  The numbers/units of likelihood are actually meaningful, and should be shown.  What is the difference between "volume" and "Gaussian"?
\end{quote}
All three figures have been reworked and are bigger. The y-axis label has been changed to "Selection Criterion." The absolute scale of these scores is not actually all that important, however it's the relative differences that encode which model is to be chosen. \hl{We might want to add more to the caption indicating that we are plotting -2logL...}

@@

We should not anthropomorphize our mathematical expressions! This original text has been removed, though the section entitled \emph{Bias-Variance Dilemma} discusses the tradeoff we were referring to in the original draft.

@@

We've removed the Gaussian likelihood results.

\separate
\begin{quote}
Fig. 3: "Since k-means optimizes an objective function that is related to the emission probability" What is this objective function?  As I understand it, k-means simply minimizes the sum of intracluster variances, which has nothing to do with your choice of emission probability being uniform or Gaussian or otherwise. Or do you mean you optimize the log-likelihood form of Eq 1 using a voronoi decomposition based on generators that are optimized for each number of states?  Also: "The likelihood decreases as new states are added..."  Which likelihoods are printed in the figure panels?  Test set, uniform, or Gaussian emission probabilities?
\end{quote}

The k-means clustering method minimizes the average mean squared distance from each data point to its cluster's mean. This is precisely the same as the $\hat{\sigma}^2$ term in the original equation. One can show that the resulting log-likelihood (evaluated on the data set) includes a term proportional to $\log \hat{\sigma}$. This is what we meant by the k-means clustering optimizing the likelihood function (and is precisely why Pelleg and Moore selected the probability that they did).

@@

\separate
\begin{quote}
``...AIC is comparable to the test set log-likelihood...The BIC penalizes complexity more strongly..." Why the huge difference between the AIC and BIC?
\end{quote}

The difference just comes from the difference between the way both are defined. The BIC penalizes the number of parameters more strongly. We've added a discussion of where the BIC and AIC come from. It's important to note that we are not attempting to advocate the use of a specific criterion, but are more interested in convincing the reader to use a statistical model selection criterion. The gold standard, of course is a cross-validation approach. Since this is not practical for most applications, we also wanted to include other selection criteria that are easier to calculate. Given our results, we would suggest the use of the AIC, but it's always important to know where these criteria come from and what assumptions are implicitly made when using one.

\separate
\begin{quote}
Fig. 4: Too small.  Missing figure title, like "Comparison of likelihood models for Fip35 WW domain trajectory data".  "Test set log-likelihoods are omitted in C as the dense matrix algebra is intractable for $k > 10^3$".  Intractable?  Really?  But this is just summing log $T_{ij}$ over N elements, isn't it?
\end{quote}

We did not intend to use the word ``intractable" in the mathematical sense. We've changed this word to ``impractical" to avoid confusion.

@@

Yes it is. For well-sampled datasets built using the state of the art methods we can still use implied timecales, however, there are many times that the timescales are not as clean and easy to interpret. We've added additional results for models built with PCA instead of tICA on the WW dataset to illustrate how the implied timescales are not necessarily  in non-ideal cases.

@@

This discussion is no longer in this revision since we've removed the Gaussian emission model to the appendix. However, in case the reviewer was curious, Lane et al. selected the number of states by insisting that each point was within 4.5 \AA of the cluster center. Kellogg et al. used a likelihood (Eq. (11) in the revision) based on a discriminative probability model.

@@

\begin{quote}
``These results take a step toward fully automating the MSM construction process...by providing a quantitative method for selecting the most suitably complex model.''  Which method in particular is recommended?  The Gaussian method?  AIC or BIC?  I'm still confused about what the actual, concrete demonstration of superiority is and which method the authors suggest should be used.
\end{quote}
\hl{how's this?}
The point that we are attempting make is to say that in many cases the implied timescale tests, which are motivated by arguments made in the limit of infinite data, are not enough for building MSMs on real data sets. It is frequently observed that the timescales are not well-behaved when we have limited data. Therefore, model selection based on timescale convergence is not able to answer the question ``given the data that I have, what is the best MSM I can build?" The point is that we need to take into account the statistical error in order to answer this question. The likelihood-based criteria are a few examples of such methods, but they may not be the gold standard for model selection. We are merely proposing that a shift in approach is needed to attempt to build models on systems that are increasingly large and that study increasingly slow timescales.

The statistical tools we've suggested match the timescale tests when the dataset is ``well-behaved" (i.e. when we have a lot of data and are using tICA). However, there are many cases where the timescales do not converge at all and so the timescale tests fail to pick a reasonable MSM. We've illustrated this with a simple example (see Fig. (4)).

\separate
\begin{quote}
Minor issues that the authors should address are below:
Abstract: ``in regimes in which these techniques encourage over-fitting''  It is unclear precisely what ``these techniques'' refers to.
\end{quote}
This is a good point, and we've rephrased to say:
\begin{quote}
Application of techniques that consider both systematic bias and statistical error on two 100$\mu s$ molecular dynamics trajectories of the Fip35 WW domain shows agreement with existing techniques based on self-consistency of the model's relaxation timescales, with more suitable results in regimes in which those timescale-based techniques encourage over-fitting.
\end{quote} 
\separate
\begin{quote}
Two key quantities which define the MSM are thus the state definitions, an indicator function basis over phase space..."  There is a substantial quantity of literature on the use of basis functions that are not indicator functions (i.e. "fuzzy" rather than "crisp" partitionings) that should be mentioned.  See, for example, the work of Marcus Weber and Susanna Roeblitz (nee Kube).
\end{quote}

@@

\begin{quote}
``but is intractable in high dimensions when the states are defined by general polytopes, such as those produced by a data-driven Voronoi tessellation''  Be clear that *exact* volume computation is intractable in very high dimension due to the complexity of all known algorithms.  Monte Carlo schemes (such as the one used here, or methods such as [http://www.cs.elte.hu/~lovasz/vol4-focs-tr.pdf]) allow the approximation of this volume.  However, it's also worth noting that the Gaussian model you propose in Eq. 4 is ALSO INTRACTABLE for EXACTLY THE SAME REASON, since you do not compute the true normalizing constant for essentially identical reasons of computational tractability in high dimension.  As such, phrases such as "A more tractable alternative is a Gaussian emission model.." are entirely disingenuous and misleading.
\end{quote}

@@

This sentence was poorly worded, but we've reworked this section, and so this should be clearer. (See emph{Results and Discussion: WW Domain}.)

@@

The reviewer is exactly right, and this is the point we're trying to drive home. It would be nice to compute the statistical error, but it is a difficult thing to calculate, since it depends on both the estimate of the transition probabilities given the state space, as well as error due to where the states are located. However, the likelihood approach will take all sources of error into account.

Reviewer 2:

@@

\hl{I have this data, so we can add it as an appendix. It doesn't hurt.}

@@

Reviewer 3:

@@

\begin{quote}
I suggest to build a fine discrete model, e.g. by direct gridding of the M\"uller potential, and constructing a simple MCMC dynamics on neighboring gridpoints, and then considering different clusterings on this data set. There, the exact timescales are known, and an embedding in a geometric space is given, such that the method can be applied such that the present method can be applied as is. It then needs to be shown that the present likelihood approach selects MSMs with timescales that coincide with the true values.In addition, it would be useful to compare the results here with implied timescale plots. Do the results agree in the cases where ITS do converge?
\end{quote}

@@

These timescales were correctly displayed. The timescales in A and B are calculated from the same set of models, but the evaluation of those models just provides a different ``score." Part B of each of these figures has been moved to the appendix.

@@

The reviewer is definitely correct here. We are not accusing anyone of ``cheating" but more drawing attention to the fact that with the absence of an optimizable objective function there is nothing preventing one from picking the model that gives a reasonable answer. \hl{we should rephrase this sentence, but I don't know what to do exactly}

\separate
\begin{quote}
Gaussians could be truncated and normalized - it seems that renormalization would require to solve an even harder problem than for uniform output probabilities, because you’d need to integrate the Gaussian density over a high-dimensional polytope.
\end{quote}

@@

\begin{quote}
I’m surprised that there are three slow relaxation processes for the Müller potential before the second gap. The potential looks like there should be only two (due to three metastable states). Were MSMs build with reversible transition matrix estimation, and are all eigenvalues thus real-valued?
\end{quote}

\begin{quote}
Muller potential: lag time was chosen to be 30 steps. How is this choice justified? Since the timescales in Fig. 2 are not independent of N, this must mean that at least for some state decompositions the ITS are not converged in the lag time. In these cases, it’s not even appropriate to describe the state-to-state dynamics by a Markov chain at all. Is the statistical model still meaningful in this case
\end{quote}

\begin{quote}
Müller potential, Gaussian likelihood: How were the Gaussian sigma-parameters chosen? The k-means states in Fig 3 have very different sizes and their centers are sometimes very close, so the assumption that output distributions shall not overlap seems to be hard to realize.
\end{quote}

\begin{quote}
WW domain, lag time was chosen to be 50 ns - same comment as for the Mueller potential, above
\end{quote}

\begin{quote}
WW domain, Gaussian likelihood - same comment as for the Müller potential, above
\end{quote}

@@

\begin{quote}
 Doesn’t Fig 4c suggest that there’s a problem here? For the likelihood-based choice made, the slowest timescale is estimated between 1500 and 3000 ns with RMSD clustering. In contrast, the TICA-based estimates suggest a timescale of around 5000 ns. So if the TICA results is indeed correct, the method should be going for larger state number in the RMSD case.
\end{quote}

\begin{quote}
In some timescale vs number of state plots the timescales descreases with increasing number of states (e.g. Fig 2A). That’s surprising. Can the authors explain that? Has a prior been used that may be responsible for this behavior?
\end{quote}

\closing{Sincerely,}

\bibliography{bibliography.bib}
\end{letter}
\end{document}

>>>>

\documentclass{article}
\usepackage{color,soul}
\usepackage{hyperref}
\usepackage{fullpage}
\usepackage{natbib}
\usepackage{cleveref}

\newcommand{\separate}{\begin{center}--------\end{center}}
\begin{document}

\hfill \begin{minipage}{10em} Clark Center S295\\ Stanford University\\ Stanford, CA, 94305 \end{minipage}

\noindent Prof. Sharon Hammes-Schiffer \\ Deputy Editor of JPC B \\ The Journal of Physical Chemistry

\vspace{2em}
\noindent Dear Prof. Hammes-Schiffer,
\vspace{1em}

We are resubmitting this manuscript, whose title is now ``Statistical Model Selection for Markov Models of Biomolecular Dynamics.'' The reviews were very helpful. Our detailed responses are included below, inline with the comments that they respond to.

\vspace{1em}
\noindent Reviewer 1:

\begin{quote}
In ``Optimal parameter selection in Markov state models for biomolecular conformational dynamics'', the authors elaborate on an earlier likelihood-based approach to the selection of the appropriate number of discrete states for crisp partitionings to attempt to obtain a Markov model that balances accuracy with the potential for overfitting that many-state models are capable of.  Three different critera are examined: a leave-some-out cross-validation likelihood estimate, as well as two criteria that penalize overfitting by using assumptions of asymptotic normality of the likelihood function (AIC and BIC).  Two different ``emission probabilities" for the probability of observing a sample $x_t$ from a given discrete state s are considered that would produce a generative model---one that can be used to assign new data to a model: a uniform probability and Gaussian probability.  These are compared for consistency with each other, as well as compared with well-established metrics such as the stability of implied timescales of the model with the number of states.
\end{quote}

@@

In order to compute the Bayes factor for selecting the number of states, we would need to integrate over the space of all crisp state decompositions $S$ with $|S|=k$ as well as the simplex of $k \times k$ reversible stochastic matrices. Although the second is possible analytically using the results of \cite{diaconis2006}, we're not aware of any practical methods for the former. To avoid confusion, we've replaced the word ``intractable'' with ``impractical.''

@@

We've expanded our discussion of the simulation methods as well as the parameters used in MSM construction to address these and other issues. The new subsection, entitled \emph{Simulation and MSM Construction} can be found in the \emph{Methods} section.

@@

\begin{quote}
``...the likelihood remained essentially constant...'' What does this mean, quantitatively?
\end{quote}

We selected $M$, the number number rounds of Monte Carlo volume estimation, such that the log likelihood was converged with respect to $M$.

\separate
\begin{quote}
``according to the three methods" Which three methods exactly?
\end{quote}
We've clarified this to read:
\begin{quote}
according to the BIC, AIC, and test log-likelihood
\end{quote}

\separate
\begin{quote}
Fig. 2: This figure is way too small.  What is ``score''?  Is this a log likelihood?  Is lower or higher score representative of lower or higher likelihood?  The numbers/units of likelihood are actually meaningful, and should be shown.  What is the difference between ``volume" and ``Gaussian"?
\end{quote}
All three figures have been reworked and are bigger. The y-axis label has been changed to ``Selection Criterion.'' As discussed by Liddle and others, ``the absolute value of the criterion is not of interest, only the relative value between different models.''\cite{liddle2004many}.

@@

This original text has been removed, though the subsection entitled \emph{Theory: Bias-Variance Dilemma} discusses the tradeoff we were referring to in the original draft.

@@

As discussed above, we've removed the Gaussian likelihood results.

\separate
\begin{quote}
Fig. 3: ``Since k-means optimizes an objective function that is related to the emission probability'' What is this objective function?  As I understand it, k-means simply minimizes the sum of intracluster variances, which has nothing to do with your choice of emission probability being uniform or Gaussian or otherwise. Or do you mean you optimize the log-likelihood form of Eq 1 using a voronoi decomposition based on generators that are optimized for each number of states?  Also: ``The likelihood decreases as new states are added...''  Which likelihoods are printed in the figure panels?  Test set, uniform, or Gaussian emission probabilities?
\end{quote}

The k-means clustering method minimizes the average mean squared distance from each data point to its cluster's mean. This is precisely the same as the $\hat{\sigma}^2$ term in the original equation. One can show that the resulting log-likelihood (evaluated on the data set) includes a term proportional to $\log \hat{\sigma}$. This is what we meant by the k-means clustering optimizing the likelihood function (and is precisely why \citet{Pelleg2000Xmeans} selected the probability that they did).

@@

\separate
\begin{quote}
``...AIC is comparable to the test set log-likelihood...The BIC penalizes complexity more strongly...'' Why the huge difference between the AIC and BIC?
\end{quote}

The difference just comes from the difference between the way both are defined. The BIC penalizes the number of parameters more strongly. We've added a discussion of the origins of the BIC and AIC. It's important to note that we are not attempting to advocate the use of a specific criterion, but are more interested in convincing the reader to use a statistical model selection criterion. The gold standard, of course is a cross-validation approach. Since this is not practical for most applications, we also wanted to include other selection criteria that are easier to calculate. Given our results, we would suggest the use of the AIC, but it's always important to know where these criteria come from and what assumptions are implicitly made when using one.

\separate
\begin{quote}
Fig. 4: Too small.  Missing figure title, like ``Comparison of likelihood models for Fip35 WW domain trajectory data".  ``Test set log-likelihoods are omitted in C as the dense matrix algebra is intractable for $k > 10^3$".  Intractable?  Really?  But this is just summing log $T_{ij}$ over N elements, isn't it?
\end{quote}

We did not intend to use the word ``intractable" in to indicate a particular computational complexity class. We've changed this word to ``impractical" to avoid confusion.

@@

Yes it is. For well-sampled datasets built using the state-of-the-art methods, the implied timescales, which consider only the systematic error in the MSM, yield consistent recommendations with the AIC, BIC and test log-likelihood. In other regimes, where the statistical error dominates, the results are different.

@@

This discussion is no longer in this revision since we've removed the Gaussian emission model. However, in case the reviewer was curious, Lane et al. selected the number of states by insisting that each point was within 4.5 \AA$\,$ of the cluster center. Kellogg et al. used a likelihood (Eq. (11) in the revision) based on a discriminative probability model.

@@

\begin{quote}
``These results take a step toward fully automating the MSM construction process...by providing a quantitative method for selecting the most suitably complex model.''  Which method in particular is recommended?  The Gaussian method?  AIC or BIC?  I'm still confused about what the actual, concrete demonstration of superiority is and which method the authors suggest should be used.
\end{quote}

Our position in this manuscript is that in many cases the implied timescale tests, which are motivated by arguments made in the limit of infinite data, are insufficient for building MSMs on available data sets. We frequently observe that the timescales are not well-behaved for many large scale simulations. Therefore, model selection based on timescale convergence is not able to answer the question ``given the data that I have, what is the best MSM I can build?" In this regime, both the statistical and systematic sources of error must be taken into account simulataneously. The likelihood-based criteria are a few examples of such methods, but they may not be the gold standard for model selection.

The statistical tools we've suggested match the timescale tests when the dataset is large and ``well-behaved.'' However, there are many cases where the timescales do not converge at all and so the timescale tests fail to pick a reasonable MSM, which we illustrate in Fig. 4.

\separate
\begin{quote}
Minor issues that the authors should address are below:
Abstract: ``in regimes in which these techniques encourage over-fitting''  It is unclear precisely what ``these techniques'' refers to.
\end{quote}
This is a good point, and we've rephrased to say:
\begin{quote}
Application of techniques that consider both systematic bias and statistical error on two 100$\mu s$ molecular dynamics trajectories of the Fip35 WW domain shows agreement with existing techniques based on self-consistency of the model's relaxation timescales, with more suitable results in regimes in which those timescale-based techniques encourage over-fitting.
\end{quote} 
\separate
\begin{quote}
``Two key quantities which define the MSM are thus the state definitions, an indicator function basis over phase space..."  There is a substantial quantity of literature on the use of basis functions that are not indicator functions (i.e. ``fuzzy'' rather than ``crisp'' partitionings) that should be mentioned.  See, for example, the work of Marcus Weber and Susanna Roeblitz (nee Kube).
\end{quote}

@@

\begin{quote}
``but is intractable in high dimensions when the states are defined by general polytopes, such as those produced by a data-driven Voronoi tessellation''  Be clear that *exact* volume computation is intractable in very high dimension due to the complexity of all known algorithms.  Monte Carlo schemes (such as the one used here, or methods such as [http://www.cs.elte.hu/~lovasz/vol4-focs-tr.pdf]) allow the approximation of this volume.  However, it's also worth noting that the Gaussian model you propose in Eq. 4 is ALSO INTRACTABLE for EXACTLY THE SAME REASON, since you do not compute the true normalizing constant for essentially identical reasons of computational tractability in high dimension.  As such, phrases such as ``A more tractable alternative is a Gaussian emission model.." are entirely disingenuous and misleading.
\end{quote}

@@

This sentence was poorly worded, but we've reworked this section, and so this should be clearer.

@@

The reviewer is exactly right, and this is precisely the point of our contribution. One advantage of the test-set likelihood approach is that it consolidates both the systematic and statistical contributions to the error into a single computable function.

\vspace{3em}
\noindent Reviewer 2:

@@

This is an interesting question, but we feel it is beyond the scope of this work. However, the approach based on likelihood functions should be able to be used to study this question.

@@

\vspace{3em}
\noindent Reviewer 3:
\vspace{1em}

Many of the comments made by Reviewer 1 and Reviewer 3 were closely related, and our responses reflect the consolidated concerns of both reviewers.

@@

These are valuable points, which we fully agree with. Our focus is on best fitting the available data, and ideally validating models on new data which was left out during training. The ultimate goal is to fit the underlying dynamics, but our only ``window'' into the underlying data is by way of MD trajectories. Clearly convergence proofs, that MSM methods converge to the exact transfer operator in the limit of infinite sampling and infinite model complexity (i.e. $k \rightarrow \infty$) are essential, but these results do not provide direct guidance to practitioners on balancing competing sources of error.

\separate
\begin{quote}
I suggest to build a fine discrete model, e.g. by direct gridding of the M\"uller potential, and constructing a simple MCMC dynamics on neighboring gridpoints, and then considering different clusterings on this data set. There, the exact timescales are known, and an embedding in a geometric space is given, such that the method can be applied such that the present method can be applied as is. It then needs to be shown that the present likelihood approach selects MSMs with timescales that coincide with the true values. In addition, it would be useful to compare the results here with implied timescale plots. Do the results agree in the cases where ITS do converge?
\end{quote}

In the finite-data regime, our methods will not select MSMs with timescales that correspond to the ``true values,'' precisely because there is a competing statistical error. Proofs of the consistency and asymptotic optimality of the AIC and BIC for linear regression do exist, which guarentee that as $n\rightarrow\infty$, the correct model will be selected with proability one\cite{nishii1984, shao1997asymptotic}. Our intuition is that similar results may apply for MSMs in the joint infinite-data / infinite number of states limit, but we have not shown this. Given that our approach is focused on practical tools in the limited data regime, we prefer to focus there.

@@

These timescales were correctly displayed. The timescales in A and B are calculated from the same set of models, but the evaluation of those models just provides a different ``score." The results with the Gaussian emission distribution have been removed.

@@

The reviewer is definitely correct here. We are not accusing anyone of ``cheating" but more drawing attention to the fact that with the absence of an optimizable objective function, a variety of human factors may come into play. We've nonetheless removed this comment from the paper.

\separate
\begin{quote}
Gaussians could be truncated and normalized - it seems that renormalization would require to solve an even harder problem than for uniform output probabilities, because you‚Äôd need to integrate the Gaussian density over a high-dimensional polytope.
\end{quote}

@@

As discussed above, we've removed the Gaussian emission distribution.

\separate
\begin{quote}
I'm surprised that there are three slow relaxation processes for the M\"{u}ller potential before the second gap. The potential looks like there should be only two (due to three metastable states). Were MSMs build with reversible transition matrix estimation, and are all eigenvalues thus real-valued?
\end{quote}

Yes, they are all real-valued.

\separate
\begin{quote}
M\"{u}ller potential: lag time was chosen to be 30 steps. How is this choice justified? Since the timescales in Fig. 2 are not independent of N, this must mean that at least for some state decompositions the ITS are not converged in the lag time. In these cases, it's not even appropriate to describe the state-to-state dynamics by a Markov chain at all. Is the statistical model still meaningful in this case?
\end{quote}

\begin{quote}
WW domain, lag time was chosen to be 50 ns - same comment as for the Mueller potential, above.
\end{quote}

Our goal is to build the best possible Markov model given the data and a prespectified lag time. From the perspective of model selection based on implied timescale convergence with respect to lag time, it's usually required that the modeller have some physical intuition about the correct number of states. Here, we work from the opposite direction, as likelihood based approach gives no guidance on the selection of $\tau$.

\separate
\begin{quote}
M\"{u}ller potential, Gaussian likelihood: How were the Gaussian sigma-parameters chosen? The k-means states in Fig 3 have very different sizes and their centers are sometimes very close, so the assumption that output distributions shall not overlap seems to be hard to realize.
\end{quote}

\begin{quote}
WW domain, Gaussian likelihood - same comment as for the M\"{u}ller potential, above.
\end{quote}

As discussed above, we've removed the Gaussian emission distribution.
\separate

@@

They were supposed to be the same timescales, but different likelihoods. Now that we have removed the Gaussian emission distribution, this point of confusion should not be present.

\separate
\begin{quote}
 Doesn't Fig 4(c) suggest that there's a problem here? For the likelihood-based choice made, the slowest timescale is estimated between 1500 and 3000 ns with RMSD clustering. In contrast, the tICA-based estimates suggest a timescale of around 5000 ns. So if the tICA results is indeed correct, the method should be going for larger state number in the RMSD case.
\end{quote}

This is a question of perspective. Our view is that even if the models based on RMSD or PCA underestimate the timescales, it nonetheless introduces an unacceptable amount of statistical error in the finite-data context to add states without limit.

\separate
\begin{quote}
In some timescale vs number of state plots the timescales descreases with increasing number of states (e.g. Fig 2A). That's surprising. Can the authors explain that? Has a prior been used that may be responsible for this behavior?
\end{quote}

We have no strong theortical justification for this observation, but we have seen slight decreases in the timescales with respect to $k$ on occation, regardless of the method by which $\mathbf{T}$ was estimated (i.e. with and without psuedocounts).

\vspace{4em}
\noindent Sincerely, \\

\hspace{5em} Robert T. McGibbon

\hspace{5em} Christian R. Schwantes

\hspace{5em} Vijay S. Pande

\bibliographystyle{plain}
\bibliography{bibliography}

\end{document}
